Lexicalized Domain Representations for Neural Machine Translation

### 2019-10-11 2019/iwslt/submission.tex
CR

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Please make sure to keep technical paper submissions anonymous  !
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\name{Firstname Lastname}
%\name{MinhQuang Pham$^{\dag\ddag}$, \ Josep Crego$^\dag$, \ Fran\c cois Yvon$^\ddag$, \ Jean Senellart$^\dag$}

@@

\address{ Affiliation / Address line 1 \\ Affiliation / Address line 2 \\ Affiliation / Address line 3 \\ {\small \tt email@domain}}
%\address{ $^\dag$SYSTRAN / 5 rue Feydeau, 75002 Paris, France \\ \small \tt firstname.lastname@systrangroup.com \\ $^\ddag$LIMSI, CNRS,  Universit\'e Paris-Saclay 91405 Orsay, France \\ {\small \tt firstname.lastname@limsi.fr}}

@@

\begin{align}
  \tilde{e}_k(v) =& M_g e_g(v) + \sum_{i \in [1,..,d]} M_i * e_i(v) * \delta(i=k) \nonumber \\
   & = M [e_g(v); e_1'(v,k) \dots; e_d'(v,k)], \label{eq:embedding}
\end{align}
where $\delta()$ is the indicator function, $M$ is the matrix made of blocks $M_g, M_1 \dots, M_d$, and $e_i'(v,k)$ is the masked embedding: $e_i'(v,k)= e_i(v) * \delta(i=k)$.

@@

Making sure that the actual embedding do not contain any zero is important for the Transformer model, since the lexical representations are then added to the positional encoding, which would undo the effect of domain masking, and propagate a gradient even to regions that should not be modified. 
With our design, we make sure that during backprogation, the matrix $M$ receives gradient $0$ at regions corresponding to deactivated regions in the word embedding. 
Those regions are also masked in forward step, thus do not interfere the training on the domains to which they are not assigned (see Figure~\ref{fig:network}).
%
Our architecture is thus readily compatible with any NMT architecture, where we simply replace standard embedding layers the computation in equation~\eqref{eq:embedding}. In our experiments, we consider both the attentional RNN architecture of \cite{Bahdanau15learning} and the Transformer architecture of \cite{Vaswani17attention}.

@@

We now evaluate the ability of our model to integrate new domains, a very common scenario for industrial MT. In this setting, we consider that we have a model ($\mathtt{LDR}_{oracle}$) trained as before for EMEA, EPPS and ECB during 200,000 iterations, which needs to process new training data from the IT domain. Assuming that we have reserved extra empty embedding cells\footnote{For this experiment, word embeddings contain 480 cells for the generic region and 32 cells for domain specific regions (8 cells x 4 regions).} for this domain, we resume training with 4 domains during 100,000 additional iterations, yielding a updated model  $\mathtt{LDR}_{oracle}^*$. Results for the English$\rightarrow$French language pair are in Table~\ref{tab:add}, where for comparison purposes we also report numbers obtained with continued training with the $\mathtt{Mixed}$ model, training for the same number of iterations and using the same four datasets ($\mathtt{Mixed}^*$).

@@

As expected, a huge improvement in performance is observed for the IT test set when learning includes in-domain data for both models, with $\mathtt{LDR}_{oracle}^*$ outperforming $\mathtt{Mixed}^*$ by a wide margin.
%
It is interesting to see that this additional data has also a positive impact on other test sets: both models similarly increase their performance for the ECB domain, and $\mathtt{LDR}_{oracle}^*$ additionally improves the results for the EMEA test, which is not the case for $\mathtt{Mixed}^*$;
finally, using IT data does not impact the quality of translations for the EPPS domain of any of the models. Overall better results are obtained by our $\mathtt{LDR}_{oracle}^*$ model trained with data from an additional source, demonstrating the ability of our architecture to seamlessly integrate new domains. % as supplementary training data becomes available.
%New domains are easily integrated in the model if the corresponding embedding regions are reserved in advance.

@@

To check this hypothesis, we performed the following analysis of embeddings learned with the multi-domain Transformer system for English:French. For each unit\footnote{In this study, we work with BPE units meaning that in many cases we observe the variation of use of \emph{word parts}. As we work with a large inventory, many of these units still correspond to actual words and we focus on these in our comments. We also restrict our analysis to words that occur at least 30 times in each domain, to ensure that each domain-specific region is updated during training.} in our English dictionary, we compute the $k$ nearest neighbours for each domain $i \in [1\dots{}d]$, where the distance between unit $u$ and $v$ for domain $i$ is the cosine distance in the corresponding embedding space, ie.\ assuming that the actual embedding of $v$ for domain $i$ is $e(v,i) = M_ge_g(u) + M_de_i(v)$ (cf.\ equation~\eqref{eq:embedding}). This process yields $d$ lists of $k$ nearest neighbours. A small intersection should then be a sign of a variation of use across domains; conversely, an near-identical set of neighbours across domains should reflect the stability of word use. Table~\ref{tab:embeddings} list the 10 units with the smaller (respectively larger) intersection (we use $k=10$ and $d=3$).

@@

The multi-domain scenario in NMT has been studied in a number of recent works. \cite{Farajian17multidomain} learn a generic system and propose to dynamically adapt the network weights in an unsupervised manner using a small sample of training data that resembles the test data, an idea already explored in \cite{Sennrich13multidomain}. Their main contribution is to propose an a method to relate the amount of adaptation of the network parameters to the similarity between the adaptation sample and the test sentence: the higher the similarity, the more agressive the adaptation.
%
By analogy with the proposal of \cite{Johnson17google} for multilingual NMT, \cite{Kobus17domaincontrol} and \cite{Chu18multilingual} separately propose to extend the representation of the source text with a domain tag. Our model also modifies input representations, but allows each source word to have a domain-specific representation, thereby improving training of the shared parts of the network.\fyDone{Talk about target embeddings somewhere.}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Please make sure to keep technical paper submissions anonymous  !
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\name{Firstname Lastname}
\name{MinhQuang Pham$^{\dag\ddag}$, \ Josep Crego$^\dag$, \ Fran\c cois Yvon$^\ddag$, \ Jean Senellart$^\dag$}

@@

%\address{ Affiliation / Address line 1 \\ Affiliation / Address line 2 \\ Affiliation / Address line 3 \\ {\small \tt email@domain}}
\address{ $^\dag$SYSTRAN / 5 rue Feydeau, 75002 Paris, France \\ \small \tt firstname.lastname@systrangroup.com \\ $^\ddag$LIMSI, CNRS,  Universit\'e Paris-Saclay 91405 Orsay, France \\ {\small \tt firstname.lastname@limsi.fr}}

@@

\begin{align}
  \tilde{e}_k(v) =& M_g e_g(v) + \sum_{i \in [1,..,d]} M_i \times e_i(v) \times \delta(i=k) \nonumber \\
   & = M [e_g(v); e_1'(v,k) \dots; e_d'(v,k)], \label{eq:embedding}
\end{align}
where $\delta()$ is the indicator function, $M$ is the matrix made of blocks $M_g, M_1 \dots, M_d$, and $e_i'(v,k)$ is the masked embedding: $e_i'(v,k)= e_i(v) * \delta(i=k)$.

@@

Making sure that the actual embedding do not contain any zero is important for the Transformer model, since the lexical representations are then added to the positional encoding, which would undo the effect of domain masking, and propagate a gradient even to regions that should not be modified. 
With our design, we make sure that during backprogation, the matrix $M$ receives gradient $0$ at regions corresponding to deactivated regions in the word embedding. 
Those regions are also masked in forward step, thus do not interfere the training on the domains to which they are not assigned (see Figure~\ref{fig:network}).
%
Our architecture is thus readily compatible with any NMT architecture, where we simply replace standard embedding layers 
by the embeddings defined
in equation~\eqref{eq:embedding}. In our experiments, we consider both the attentional RNN architecture of \cite{Bahdanau15learning} and the Transformer architecture of \cite{Vaswani17attention}.

@@

We now evaluate the ability of our model to integrate new domains, a very common scenario for industrial MT. In this setting, we consider that we have a model ($\mathtt{LDR}_{oracle}$) trained as before for EMEA, EPPS and ECB during 200,000 iterations, which needs to process new training data from the IT domain. Assuming that we have reserved extra empty embedding cells\footnote{For this experiment, word embeddings contain 480 cells for the generic region and 32 cells for domain specific regions (8 cells x 4 regions).} for this domain, we resume training with 4 domains during 100,000 additional iterations, yielding an updated model  $\mathtt{LDR}_{oracle}^*$. Results for the English$\rightarrow$French language pair are in Table~\ref{tab:add}, where for comparison purposes we also report numbers obtained with continued training with the $\mathtt{Mixed}$ model, training for the same number of iterations and using the same four datasets ($\mathtt{Mixed}^*$).

@@

As expected, a huge improvement in performance is observed for the IT test set when learning includes in-domain data for both models, with $\mathtt{LDR}_{oracle}^*$ outperforming $\mathtt{Mixed}^*$ by a wide margin.
%
It is interesting to see that this additional data has also a positive impact on other test sets: both models similarly increase their performance for the ECB domain, and $\mathtt{LDR}_{oracle}^*$ additionally improves the results for the EMEA test, which is not the case for $\mathtt{Mixed}^*$;
finally, using IT data does not impact the quality of translations for the EPPS domain of any of the models. Overall better results are obtained by our $\mathtt{LDR}_{oracle}^*$ model trained with data from an additional source, demonstrating the ability of our architecture to seamlessly integrate a new domain. % as supplementary training data becomes available.
%New domains are easily integrated in the model if the corresponding embedding regions are reserved in advance.

@@

To check this hypothesis, we performed the following analysis of embeddings learned with the multi-domain Transformer system for English:French. For each unit\footnote{In this study, we work with BPE units meaning that in many cases we observe the variation of use of \emph{word parts}. As we work with a large inventory, many of these units still correspond to actual words and we focus on these in our comments. We also restrict our analysis to words that occur at least 30 times in each domain, to ensure that each domain-specific region is updated during training.} in our English dictionary, we compute the $k$ nearest neighbours for each domain $i \in [1\dots{}d]$, where the distance between unit $u$ and $v$ for domain $i$ is the cosine distance in the corresponding embedding space, ie.\ assuming that the actual embedding of $v$ for domain $i$ is $e(v,i) = M_ge_g(u) + M_ie_i(v)$ (cf.\ equation~\eqref{eq:embedding}). This process yields $d$ lists of $k$ nearest neighbours. A small intersection should then be a sign of a variation of use across domains; conversely, an near-identical set of neighbours across domains should reflect the stability of word use. Table~\ref{tab:embeddings} list the 10 units with the smaller (respectively larger) intersection (we use $k=10$ and $d=3$).

@@

The multi-domain scenario in NMT has been studied in a number of recent works. \cite{Farajian17multidomain} learn a generic system and propose to dynamically adapt the network weights in an unsupervised manner using a small sample of training data that resembles the test data, an idea already explored in \cite{Sennrich13multidomain}. Their main contribution is to propose a method to relate the amount of adaptation of the network parameters to the similarity between the adaptation sample and the test sentence: the higher the similarity, the more agressive the adaptation.
%
By analogy with the proposal of \cite{Johnson17google} for multilingual NMT, \cite{Kobus17domaincontrol} and \cite{Chu18multilingual} separately propose to extend the representation of the source text with a domain tag. Our model also modifies input representations, but allows each source word to have a domain-specific representation, thereby improving training of the shared parts of the network.\fyDone{Talk about target embeddings somewhere.}
